# Methods

```{r table-of-events, eval=FALSE}
events <- readRDS("data/events/england.Rds") %>% sample_n(20000)

all_events <- events %>% 
  tbl_df() %>% 
  ungroup() %>% 
  select(event_name, sub_event_name) %>% 
  group_by(event_name, sub_event_name) %>% 
  distinct() %>% 
  arrange(event_name, sub_event_name) 

all_events %>% 
  filter(event_name < "Goalkeeper") %>% 
  kable(col.names = c("Event Name", "Sub-Event Name")) %>% 
  collapse_rows(columns = 1, valign = "top") %>%
  kable_styling(full_width = FALSE, position = "float_left", fixed_thead = T)
all_events %>% 
  filter(event_name >="Goalkeeper") %>% 
  kable(col.names = c("Event Name", "Sub-Event Name")) %>% 
  collapse_rows(columns = 1, valign = "top") %>%
  kable_styling(full_width = FALSE, position = "left", fixed_thead = T) 
```

```{r making-polygrids-zones-plot}
# p1 <- grids5x5 %>% 
#   mutate(rowid = 1:294) %>% 
#   ggplot() + 
#   fc_annotate_pitch(fill = NA,color = "#F8FAF7") +
#   theme_void() + theme(plot.background = element_rect(fill = "#629A52", 
#         color = NA), aspect.ratio = 70/105) + 
#   geom_sf(fill = NA) + 
#   geom_sf_text(aes(label = rowid)) 
#   
# p2 <- zones %>%  
#   mutate(rowid = 1:8) %>% 
#   ggplot() +
#   fc_annotate_pitch(fill = "NA",color = "#F8FAF7") +
#   theme_void() + theme(plot.background = element_rect(fill = "#629A52", 
#         color = NA), aspect.ratio = 70/105, legend.position = "none")  + 
#   geom_sf(fill = NA) + 
#   #geom_sf(aes(fill = as.factor(rowid)), color = NA, alpha = 0.8) + 
#   geom_sf_label(aes(label = rowid)) 
#   # geom_sf_label(aes(label = rowid),
  #               nudge_x = c(-0.6,-2.75,-2,-2.75,2.75,2,2.75,0.6),
  #               nudge_y = c(2.5,0,2.5,0,0,2.5,0,2.5)) + 
  #geom_sf(data = grids5x5, aes(geometry = geometry), fill=NA, color = "#2E2E2E50") + 
  #fc_annotate_arrow(x = 52.5, y = -10, palette = "bw") + 
  
#gridExtra::grid.arrange(p1,p2, nrow=1)
```

## Possession ID 

Ball possession is broadly defined as the amount of time a team possesses the ball during a game. However, there is no widely accepted benchmark/definition of what events concludes a possession and triggers a new one [[6]][References]. Thus, we created a possession identifier that indicates the current unique possession in a game, and we defined new possessions as starting after a team demonstrates that it has established control of the ball. This occurs in the following situations: at the start of a half, when the team intercepts or successfully tackles the ball, and after the opposing team last touches the ball before it goes out of play, commits a foul followed by a freekick, or after a shot is taken. A new possession can also begin even if the same team has possession of the ball. For example, when the ball goes out for a throw for the attacking team, we have a new possession for the same attacking team. In addition, if the attacking team makes a pass after a sequence of duels, events in which opposing players contest the ball, this constitutes the same possession. According to our definition above, there were an average of 306 possessions per game. 

## Metrics of Pace 

**[[OMA: Ethan, move this first sentence to the third paragraph where you define them.]]**
After creating a possession identifier, we calculated the total ($V_{T}$), east-west ($V_{EW}$), north-south ($V_{NS}$), and east-only ($V_{E}$) velocities of each event.  We first calculate distance traveled for each event. The east-west distances ($\delta_{EW}$) are determined by the difference of the starting and ending x-coordinates while the north-south distances ($\delta_{NS}$) are determined by the difference of the starting and ending y-coordinates. The total distances ($\delta_{T}$) are then calculated as $\sqrt{(\delta_{EW}^2 + \delta_{NS}^2)}$. Events are assigned an E-only distance ($\delta_{E}$) only if the pass travels toward the opposing goal. Although the ball rarely travels in a straight line, the dataset does not provide information about its trajectory, so we assume that the ball travels in a straight line from its starting to ending coordinates. There is also no indication if a player dribbles the ball before making a pass. 

Next, we determined the duration between events. For each event, the dataset only provides a timestamp in seconds since the beginning of the current half of the game. Thus, within each possession, the duration for an event was calculated as the difference of the timestamp of the following event and that of the current event.  With this definition of duration, the last event in the possession sequence is not included in the calculation of pace. Finally, the speeds of each event are calculated by dividing the corresponding distance by its duration. 

We used the distance travelled and duration between successive passes and free kicks (except for free kick shots and penalty kicks) **[[OMA: Ethan, you also make this clear in the first line of the next paragraph. You can delete it here.]]** during the same possession to calculate four different measures of pace, which include $V_{T}$, $V_{EW}$, $V_{NS}$, and $V_{E}$. $V_{E}$ differs from $V_{EW}$ in that only forward progress is measured, and any backward progress is excluded from the analysis. These metrics are the average velocities of the event rather than the instantaneous velocities, since we did not have access to tracking data. 

When analyzing pace, we only included passes and free kicks (except for free kick shots and penalty kicks) since these events are reliable indicators of the pace of the game. In addition, we only kept possessions that consist of three or more pass or free kick events, as these types of possessions are more definitive of a team’s pace.(MOVE THIS EARLIER???). **[[OMA:  Yes.]]** For the following results and discussion of pace, events will only refer to these passes and free kicks. Following the exclusion of certain possessions, the remaining possessions contained an average of 5.5 events per possession. 

All of the previously mentioned procedures can be implemented with functions in the [**scoutr**](https://github.com/shawnsanto/scoutr) package. 

## Spatial Polygrid Analysis 

We rescaled the pitch to 105 by 70 meters instead of 105 by 68 meters, to be able to divide the pitch into 294 equal, non-overlapping 5x5 meter square polygrids [[7]][References]. $V_{T}$, $V_{EW}$, $V_{NS}$, and $V_{E}$ for a given event was assigned to all polygrids that the event's path intersects. Each polygrid contains $n$ velocity values for the $n$ event paths that intersect it. For each of the 5x5 polygrids, we then take the median for each of the four different pace metrics. There are polygrids, particularly ones in the corners or along the attacking team's goal line, that have very few recorded velocity values because only a few events intersect those polygrids. These polygrids often contain passes with extremely high velocities, most of which are due to tagging errors. Thus, the median was taken, instead of the mean, to account for the presence of outliers.

## Zonal Analysis 

We divided the pitch into 8 regions. For each zone, we determined which of the 294 5x5 polygrids intersect the zone. As seen in Figure 3.1, there are some polygrids that fall into multiple zones. We then take the mean of the median $V_{T}$, $V_{EW}$, $V_{NS}$, and $V_{E}$ values of those 5x5 polygrids to determine the aggregate speeds for the zone. (**should i say there is not a big difference between taking mean of the medians vs median of the median**) **[[OMA:  Yes briefly.]]** 

This method was conducted in favor of another one that assigns an event’s speeds to all zones that intersect the path of the event. Our approach automatically factors in the event’s distance within the zone and is more resistant to outliers. For example, for a pass that intersects $n$ different 5x5 polygrids in a zone, the zone’s aggregate speed will be affected by that pass’ speed $n$ times instead of just once. This approach is thus more resistant to outliers. 

```{r fig.align = "center", fig.cap="Plot of the 294 polygrids and 8 zones overlaid on the pitch. The grey lines represent the polygrids and black borders represent the boundaries of the 8 zones."}
zones %>%  
  mutate(rowid = 1:8) %>% 
  ggplot() +
  fc_annotate_pitch(fill = "NA",color = "#F8FAF7") +
  theme_void() + theme(plot.background = element_rect(fill = "#629A52", 
        color = NA), aspect.ratio = 70/105, legend.position = "none")  + 
  
  geom_sf(color = "black", alpha = 0.8, size = 0.7) + # border around region, lighter color for polygrids 
  #geom_sf_label(aes(label = rowid)) 
  geom_sf_label(aes(label = rowid),
                nudge_x = c(-0.6,-2.75,-2,-2.75,2.75,2,2.75,0.6),
                nudge_y = c(2.5,0,2.5,0,0,2.5,0,2.5),size = 5) +
  geom_sf(data = grids5x5, aes(geometry = geometry), fill=NA, color = "#2E2E2E35")
```

## Modeling

After creating the pace-of-play metrics, we wanted to evaluate their effectiveness when used as variables in models that predict the outcome of a game. We implemented models with and without the pace metrics to determine if the models with the pace metrics achieve a higher accuracy. Our main goal is to create a predictive model that teams can use to predict the outcome of the game before the actual game. Therefore, we only considered pre-game variables that can be gathered before each game is played. **[[OMA: Ethan, talk about the fact that even though we use pace metrics from the same games, we envisage that teams can make those pre-game by substituting what we have done for historical measurements of the pace metrics.]]** Thus, when selecting variables, we did not include traditional post-game performance-based features, such as number of shots or corners taken. **I can't think of a good reason why we don't include peformance based stats** **[[OMA: with the new updates, you should be good.]]**

- Pace is something that can be modeled, but this is future work. --> tie this into writeup

```{r}
# https://stackoverflow.com/questions/53394214/r-kableextra-trouble-inserting-linebreak-in-html-format-in-a-cell-containing/53394439
# vars <- c("FTR", "HomeWinStreak", "HomeUnbeatenStreak", "HomeTeam", "DerbyGame",  "League", "$\\Delta_{ij}$Points1617")
# var_descriptions <- linebreak(c("Difference in game points from 2016-17 season (home - away)", "Outcome of a game", "Longest home win streak from 2016-17 season", "Longest home unbeaten streak from 2016-17 season", "Name of the home team", "Indicates if game is a derby",  "League that game takes place in"))
# var_values <- c(
#   "-69 to 69",
#   "Win, Draw, Loss",
#   "1 to 17",
#   "2 to 19",
#   "Manchester City, Barcelona, etc. ",
#   "0 (No), 1 (Yes)",
#   "EPL, Ligue 1, Bundesliga, Serie A, La Liga"
# )

vars <- c("FTR", "CurrentUnbeatenStreak", "DerbyGame",  "League", "HomeTeam")
var_descriptions <- linebreak(c("Outcome of a game with respect to home team", "Current home unbeaten streak (win/draw)", "Indicates if game is a derby",  "League that game takes place in", "Name of the home team"))
var_values <- c(
  "1 (Win), 0 (Draw), -1 (Loss)",
  "0 to 19",
  "0 (No), 1 (Yes)",
  "EPL, Ligue 1, Bundesliga, Serie A, La Liga",
  "Manchester City, Barcelona, etc."
)

tibble(
  vars = vars,
  var_descriptions = var_descriptions,
  var_values = var_values
)  %>%
  kableExtra::kbl(caption = "Description of Variables",
        col.names = c("Variable", "Description", "Values"),
        booktabs = TRUE, align = "lcc") %>%
  kableExtra::kable_styling()  %>%
  kableExtra::collapse_rows(columns = c(1,3))  %>%
  kableExtra::pack_rows("Response", 1, 1) %>%
  kableExtra::pack_rows("Predictors", 2,5)
  
```

For each of the 1826 games, the response variable, *FTR*, describes the outcome with respect to the home team. A *FTR* of 1 indicates that the home team won while -1 indicates the away team won. *CurrentUnbeatenStreak* is the home team's current unbeaten streak leading up to a game. The streak resets to 0 when a team loses at home. For the first home game of the season, a team's unbeaten streak from the 2016-17 season is used. For example, Manchester City went unbeaten in their last 12 home games in the 2016-17 season, so their *CurrentUnbeatenStreak* for their first home game is 12. For the 14 newly promoted teams, their *CurrentUnbeatenStreak* for their first home game is 0. This is because they played in a lower division, so their home unbeaten streak is not comparable to that of a team who played in the first division. *DerbyGame* is an indicator variable, where 1 indicates whether or not a game is a derby game. A game is marked as a derby game if the two teams are located in the same city (Manchester City vs. Manchester United) or if there is a historical rivalry (El Clásico). *League* specifies which of the five leagues the game takes place in and *HomeTeam* is the name of the home team. **does league even need to be in the table if i'm not using it in the models? ** **[[OMA: Yes. Explain why we don't use league in the model selection process.]]**

```{r}
# https://stackoverflow.com/questions/53394214/r-kableextra-trouble-inserting-linebreak-in-html-format-in-a-cell-containing/53394439
pace_vars <- c("$\\Delta_{ij}$AllZones", "$\\Delta_{ij}$OffZones", "$\\Delta_{ij}$FlankZones")
pace_var_descriptions <- linebreak(c("Sum of the differences in total velocity for all zones (1-8) for home team $i$ and away team $j$",
                                     "Sum of the differences in total velocity for all zones in offensive half (5-8) for home team $i$ and away team $j$",
                                     "Sum of the differences in total velocity for zones 5,7,8 for home team $i$ and away team $j$"))
 
pace_var_values <- c(
  "-88.93 to 67.72 (m/s)",
  "-80.68 to 45.55 (m/s)",
  "-79.1 to 39.74 (m/s)"
)

tibble(
  vars = pace_vars,
  var_descriptions = pace_var_descriptions,
  var_values = pace_var_values
)  %>%
  kableExtra::kbl(caption = "Description of Pace Variables",
        col.names = c("Variable", "Description", "Values"),
        booktabs = TRUE, align = "lcc") %>%
  kableExtra::kable_styling()  %>%
  kableExtra::collapse_rows(columns = c(1,3))  %>%
  kableExtra::column_spec(1, width = "13em") %>%
  kableExtra::column_spec(2, width = "27.5em")
```

For each game, we conducted a zonal analysis (hyperlink to previous section or something?) of the total velocities for the home and away teams. However, we took the median of the median velocities of the 5x5 polygrids to determine the aggregate speeds for each zone instead of the mean of the medians. Bottom-tier teams have a smaller number of recorded events per game and are more susceptible to outliers in both the polygrid and zonal analyses. Thus, using the med ian of the median velocities makes these zonal speeds more resistant to outliers. Then we calculated the difference (home - away) in velocity for each of the 8 zones. For home team $i$ and away team $j$, $\Delta_{ij} AllZones$ is the sum of the differences for all 8 zones, $\Delta_{ij} OffZones$ is the sum of the differences for the four zones in the offensive half (zones 5-8) and $\Delta_{ij} FlankZones$ is the sum of the differences for zones 5, 7, 8. 

**[[OMA: It might be better to change the naming conventions a bit just to make things look less clunky. How about something like $\Delta_{ij}^{FK}$, $\Delta_{ij}^{Flank}$ or even $\Delta_{ij}^{5,7,8}$ instead of $\Delta_{ij} FlankZones$? You can try the same for the other notations and even the variable names.]]**

### Model Selection 

To evaluate the models, we first split the data into a train and test set. The test data includes 2 home and 2 away games for each of the 98 teams, for a total of 392 games **[[OMA: Ethan, what percentage of the full data is this? 20% right? Add that.]]**. We perform 4-fold cross validation on the training data and lastly assess model performance by predicting on the testing data. We propose two different types of models - the first is a hierarchical logistic regression model that predicts between wins and non-wins (draws or losses) while second is a multinomial logistic regression model that predicts on all three potential outcomes. These models were preferred over other classification algorithms since we are concerned with both predictive power and interpretability. For both sets of models, the baseline model used predictors mentioned in Table 3.1. We then added one of the pace variables from Table 3.2 to determine if the addition of a pace metric improves the model's accuracy. Only one pace metric can be added to the model since they are all highly correlated. Interaction effects between the baseline predictors were also considered, but they did not improve the predictive power of any model. We used accuracy and AUC (**do i need to define auc/TPR?** **[[OMA: I don't think so.]]**) as evaluation metrics for the hierarchical logistic regression and accuracy and True Positive Rate (TPR) for the multinomial logistic regression. The final baseline and pace models are displayed below. **how to select "final" pace model - highest train accuracy or test accuracy?** **[[OMA: The results from both are consistent correct?]]**

**Shawn, Michael - thoughts on the notation?**


**[[OMA: The baseline logistic model (without any pace metrics) is as follows:]]**
$$ 
\begin{aligned}
Y_{ij} \sim Bernoulli(P_{ij}&) \\
log(\frac{P_{ij}}{1-P_{ij}}) = \alpha_{k} + \beta_{0} + \beta_{1} * Current&UnbeatenStreak_i \ + \\
\beta_{2} *  I(DerbyGame_{ij}=&Yes)  & (1.1)\\ \\ 
\alpha_k \sim N(0,\tau_0^2) 
\end{aligned}
$$

**[[OMA: The modified logistic model (with the pace metric) is as follows:]]**

$$ 
\begin{aligned}
\\\\
Y_{ij} \sim Bernoulli(P_{ij}&) \\\\
log(\frac{P_{ij}}{1-P_{ij}}) = \alpha_{k} + \beta_{0} + \beta_{1} * Current&UnbeatenStreak_i \ + \\
\beta_{2} *  I(DerbyGame_{ij}=Yes) + \beta_3& * \Delta_{ij}FlankZones & (1.2)\\ 
\alpha_k \sim N(0,\tau_0^2) 
\end{aligned}
$$

<!-- The final logistic hierarchical models are shown in 1.1 and 1.2.  -->

Here, $Y_{ij}$ is the outcome (win or loss) of a game in league $k$ (is $k$ necessary to add to model/variables? **[[OMA: No. Team is your hierarchical variable right?]]** ) when home team $i$ plays away team $j$ and $P_{ij}$ is the probability that home team $i$ defeats away team $j$. $\alpha_k$ represents the random intercept term for *HomeTeam* $k$. The only difference between models 1.1 and 1.2 is the addition of the variable $\Delta_{ij}OffZones$ in 1.2. **[[OMA: Don't you mean $\Delta_{ij}FlankZones$?]]** 

**[[OMA: Let's go over the notation more carefully on Friday.]]** 

$P_{ij\_k}$or $P_{ij\_outcome}$? **Not formating latex until notation looks good**

$$
\begin{aligned}
log(\frac{P_{ij\_k}}{1-P_{ij\_win}}) = \beta_{0k} + \beta_{1k} * CurrentUnbeatenStreak_i + \ \\
\beta_{2k}* I(DerbyGame_{ij}=Yes)  + \sum_{T = 2}^{98}\beta_{3kT}*I(HomeTeam_{i}=T) &&(2.1)\\ 
k \in  (draw, loss)
\end{aligned}
$$

$$
\begin{aligned}
\\\\
log(\frac{P_{ij\_k}}{1-P_{ij\_win}}) = \beta_{0k} + \beta_{1k} * CurrentUnbeatenStreak_i + \ \\
\beta_{2k}* I(DerbyGame_{ij}=Yes)  + \sum_{T = 2}^{98}\beta_{3kT}*I(HomeTeam_{i}=T) + \beta_{4k} * \Delta_{ij}FlankZones &&(2.2)\\ 
k \in  (draw, loss)
\end{aligned}
$$

In the multinomial regression, we set the baseline of the response variable $FTR$ as a win. $P_{ijk}$ is the probability that the game ends in a draw when $k = draw$ and a loss when $k = loss$. For the term $HomeTeam$, $T=1$ is the baseline, which is Manchester City.  The only difference between models 2.1 and 2.2 is the addition of the variable $\Delta_{ij}OffZones$ in 1.2.


