--- 
title: "Analyzing Pace-of-Play in Soccer using Spatio-Temporal Event Data"
author: "Ethan Shen"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
link-citations: yes
knit: "bookdown::render_book"
space_between_paragraphs: true
classoption: oneside
fontsize: 12pt
geometry: "left=3cm, right=3cm, top=2.5cm, bottom=2.5cm"
header-includes:
  - \usepackage{floatrow}
  - \usepackage{graphicx}
  - \usepackage{hyperref}
  - \usepackage{caption}
  - \captionsetup[figure]{font=footnotesize}
  - \captionsetup[table]{font=footnotesize}
---



```{r include=FALSE}

# STUFF IN YAML 
# bibliography: [book.bib, packages.bib]
# biblio-style: apalike
#description: "insert description."



# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

```{r include_packages, include = FALSE}
# This chunk ensures that the thesisdowndss package is
# installed and loaded. This thesisdowndss package includes
# the template files for the thesis.
if(!require(devtools))
  install.packages("devtools", repos = "http://cran.rstudio.com")
if(!require(thesisdowndss))
  devtools::install_github("mine-cetinkaya-rundel/thesisdowndss")
library(thesisdowndss)
library(knitr)
```

# Abstract {-}

Pace-of-play is an important characteristic in soccer that can influence the style and outcome of a match. Using event data provided by Wyscout that covers one season of regular-season games from five European soccer leagues, we develop four velocity-based pace metrics and examine how pace varies across the pitch, between different leagues, and between different teams. Our findings show that although pace varies considerably, it is generally highest in the offensive third of the pitch, relatively consistent across leagues, and increases with decreasing team quality. In addition, we assess whether the pace metrics are useful in predicting the outcome of a match by constructing models with and without the metrics. After implementing hierarchical logistic and multinomial logistic models, we find that the pace variables are statistically significant but only slightly improve the predictive accuracy metrics.

*Keywords*: Soccer analytics, pace-of-play, pass velocities, classification algorithms


## Acknowledgements {-}

insert this later 


