--- 
title: "Analyzing Pace-of-Play in Soccer using Spatio-Temporal Event Data"
author: "Ethan Shen"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
description: "insert description."
knit: "bookdown::render_book"
sspace_between_paragraphs: true
url: 'https\://bookdown.org/john/awesome/'
header-includes:
  - \usepackage{floatrow}

---



```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

```{r include_packages, include = FALSE}
# This chunk ensures that the thesisdowndss package is
# installed and loaded. This thesisdowndss package includes
# the template files for the thesis.
if(!require(devtools))
  install.packages("devtools", repos = "http://cran.rstudio.com")
if(!require(thesisdowndss))
  devtools::install_github("mine-cetinkaya-rundel/thesisdowndss")
library(thesisdowndss)
library(knitr)
```



# Abstract {-}

Pace-of-play is an important characteristic in soccer that can influence the style and outcome of a match. Using event data provided by Wyscout that covers one season of regular-season games from five European soccer leagues, we developed four velocity-based pace metrics and analyzed how pace varies across the pitch, between different leagues and different teams. Our findings showed that although pace varies considerably, it is generally highest in the offensive third, relatively consistent across leagues, and increases as a team’s quality decreases / displays a negative correlation with a team’s quality. We then assessed whether the pace metrics are useful in predicting the outcome of a match by constructing models with and without the metrics. After leveraging hierarchical logistic and multinomial logistic models, we found that even though the pace metrics are significant when predicting match outcomes, they only slightly improve the model’s accuracy. 

## Acknowledgements {-}

insert this if necessary lol


